---
title: "Other Analysis"
author: "Matthew Vu"
date: "2023-02-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Data
The dataset is used for this project is titled "diabetes_binary_5050split_health_indicators_BRFSS2015" by kaggle user Alex Teboul.  The dataset is a 6.35 megabyte .csv file, and is derived from the CDC's the CDC's Behavioral Risk Factor Surveillance System (BRFSS), which is the United State's most reliable, documented, and researched system for health-related telephone surveys for all 50 states and territories.  Also, the dataset consists of 70,692 observations and 22 variables.  These 70,692 observations from the 2015 BRFSS survey responses; however, the distributor of the dataset  to the These 22 variables pertain to answers to health-related survey questions, such as general health status, habits, behaviors, exercise, and chronic health conditions, as well as  calculated values based on these answers, for instance, body mass index (BMI).

## Load in data
The code below loads in the data
```{r}
data = read.csv("data.csv", 
                header = T, 
                colClasses = c("factor", # Diabetes_binary
                               "factor", # HighBP
                               "factor", # HighChol 
                               "factor", # CholCheck,
                               "numeric", # BMI
                               "factor", # Smoker
                               "factor", # Stroke
                               "factor", # HeartDiseaseorAttack
                               "factor", # PhysActivity
                               "factor", # fruits
                               "factor", # veggies
                               "factor", # HvyAlcoholConsump
                               "factor", # AnyHealthcare,
                               "factor", # NoDocdbCost,
                               "factor", # GenHlth
                               "numeric", # MentHlth
                               "numeric", # PhysHlth
                               "factor", # DiffWalk
                               "factor", # Sex,
                               "numeric", # Age,
                               "factor", # Education
                               "factor" # Income
                               ))
```

## split the data into training and test set
The code below performs the data split
```{r}
set.seed(123)
sample = sample(c(TRUE, FALSE), nrow(data), replace=TRUE, prob=c(0.7,0.3))
train_data = as.data.frame(data[sample, ])
test_data = as.data.frame(data[!sample, ])
```

# Perform exploratory data analysis

## First, analyze smoking

The code below displays the distribution of smokers in the dataset
```{r}
train_data %>%
  ggplot() + 
  geom_bar(mapping = aes(x = Smoker)) +
  labs(title = "The distribution of smokers in the dataset", 
  subtitle = "Have you smoked at least 100 cigarettes in your entire life?") + 
  scale_x_discrete(labels = c(FALSE, TRUE)) +
  theme_classic()
```
The plot above shows that the dataset mostly consists of individuals who are non-smokers.

The code below draws the graph for the proportion of of diabetic and non-diabetic people who are smokers and non-smokers. 
```{r}
train_data %>%
ggplot() +
  geom_bar(mapping = aes(x = Diabetes_binary, fill = Smoker)) +
  labs(title = "The proportion of diabetic and non-diabetic people who 
       smoked more than 100 cigarettes in their life", 
       x = "Diabetes") +
  scale_x_discrete(labels = c(FALSE, TRUE)) + 
  scale_fill_discrete(labels = c(FALSE, TRUE)) +
  theme_classic()
```
The plot above shows that a greater proportion of people who have diabetes have smoked more than 100 cigarettes in their life.

The code below performs a chi-square test of independence to see if the diabetes and smoking are independent or related.
```{r}
chisq.test(table(train_data$Smoker, train_data$Diabetes_binary))
```
Because the p-value is less than 0.05, we have sufficient evidence that diabetes and smoking more than 100 cigarettes in ones life are significantly associated.

## Second, analyize high blood pressure

The code below displays the distribution of individuals with high blood pressure in the dataset
```{r}
train_data %>%
  ggplot() + 
  geom_bar(mapping = aes(x = HighBP)) +
  labs(title = "The distribution of individuals with high blood pressure in the dataset", 
       subtitle = "Adults who have been told they have high blood pressure by a doctor, nurse, or other health professional") + 
  scale_x_discrete(labels = c(FALSE, TRUE)) +
  theme_classic()
```
The plot above shows that the dataset has a greater proportion of individuals who have high blood pressure than do not.

The code below draws the graph for the proportion of of diabetic and non-diabetic people who have and do not have high blood pressure. 
```{r}
train_data %>%
ggplot() +
  geom_bar(mapping = aes(x = Diabetes_binary, fill = HighBP)) +
  labs(title = "The proportion of diabetic and non-diabetic 
       people who have and do not have high blood pressure", 
       x = "Diabetes") +
  scale_x_discrete(labels = c(FALSE, TRUE)) + 
  scale_fill_discrete(labels = c(FALSE, TRUE)) +
  theme_classic()
```
The plot above shows that a greater proportion of people who have diabetes have high blood pressure.


The code below performs a chi-square test of independence to see if the diabetes and high blood pressure are independent or related.
```{r}
chisq.test(table(train_data$HighBP, train_data$Diabetes_binary))
```
Because the p-value is less than 0.05, we have sufficient evidence that diabetes and high blood pressure are significantly associated.

## Third, analyze difficulty walking
The code below displays the distribution of individuals who have difficulty walking or climbing stairs in the dataset
```{r}
train_data %>%
  ggplot() + 
  geom_bar(mapping = aes(x = DiffWalk)) +
  labs(title = "The distribution of individuals who have difficulty walking or climbing stairs in the dataset", 
       subtitle = "Do you have serious difficulty walking or climbing stairs?") + 
  scale_x_discrete(labels = c(FALSE, TRUE)) +
  theme_classic()
```
The plot above shows that the dataset has a greater proportion of individuals do not have difficulty walking or climbing stairs.

The code below draws the graph for the proportion of of diabetic and non-diabetic people who have difficulty walking or climbing stairs. 
```{r}
train_data %>%
ggplot() +
  geom_bar(mapping = aes(x = Diabetes_binary, fill = DiffWalk)) +
  labs(title = "The proportion of diabetic and non-diabetic 
       people who have difficulty walking or climbing stairs", 
       x = "Diabetes") +
  scale_x_discrete(labels = c(FALSE, TRUE)) + 
  scale_fill_discrete(labels = c(FALSE, TRUE)) +
  theme_classic()
```
The plot above shows that a greater proportion of people who have diabetes have difficulty walking or climbing stairs.


The code below performs a chi-square test of independence to see if the diabetes and diffuclty walking or climbing stairs are independent or related.
```{r}
chisq.test(table(train_data$DiffWalk, train_data$Diabetes_binary))
```
Because the p-value is less than 0.05, we have sufficient evidence that diabetes and difficulty walking or climbing stairs are significantly associated.

## Fourth, analyze heavy drinkers

The code below displays the distribution of heavy drinkers in the dataset
```{r}
train_data %>%
  ggplot() + 
  geom_bar(mapping = aes(x = HvyAlcoholConsump)) +
  labs(title = "The distribution of heavy drinkers in the dataset", 
  subtitle = "Adult men having more than 14 drinks per week and adult women having more than 7 drinks per week") + 
  scale_x_discrete(labels = c(FALSE, TRUE)) +
  theme_classic()
```
The plot above shows that the dataset mostly consists of individuals who are not heavy drinkers.

The code below draws the graph for the proportion of of diabetic and non-diabetic people who are smokers and non-smokers. 
```{r}
train_data %>%
ggplot() +
  geom_bar(mapping = aes(x = Diabetes_binary, fill = HvyAlcoholConsump)) +
  labs(title = "The proportion of diabetic and non-diabetic people who 
       are heavy drinkers", 
       x = "Diabetes") +
  scale_x_discrete(labels = c(FALSE, TRUE)) + 
  scale_fill_discrete(labels = c(FALSE, TRUE)) +
  theme_classic()
```
The plot above shows that a greater proportion of people who do not have diabetes are heavy drinkers.

The code below performs a chi-square test of independence to see if the diabetes and heavy drinking are independent or related.
```{r}
chisq.test(table(train_data$HvyAlcoholConsump, train_data$Diabetes_binary))
```
Because the p-value is less than 0.05, we have sufficient evidence that not have diabetes and heavy drinking are significantly associated.

This is strange because alcohol has a lot of sugar, which can equate to diabetes.  However, it makes sense that there are a greater proportion of non-diabetic people who are heavy drinkers than diabetic people because when people have diabetes, they do not want to drink because it might kill them.  As a result, this variable might not be the best variable to continue on with our analysis.  Also, this might be caused by the fact that so few observations are heavy drinkers.  A better variable might be "have you been a heavy or binge drinker in the past".

## Perform hypothesis testing of diet

The code below displays the distribution of people who eat fruits in the dataset
```{r}
train_data %>%
  ggplot() + 
  geom_bar(mapping = aes(x = HvyAlcoholConsump)) +
  labs(title = "The distribution of heavy drinkers in the dataset", 
  subtitle = "Adult men having more than 14 drinks per week and adult women having more than 7 drinks per week") + 
  scale_x_discrete(labels = c(FALSE, TRUE)) +
  theme_classic()
```




